---
title: "Lab 4: Does Prenatal Care Improve Infant Health?"
author: "w203: Statistics for Data Science"

date: "April 10, 2017"
output: pdf_document
---

# Introduction

This is a group lab.  You may work in teams of 2 or 3.

The file bwght\_w203.RData contains data from the National Center for Health Statistics and from birth certificates.  Your team has been hired by a health advocacy group to study this data and help them understand whether prenatal care improves health outcomes for newborn infants.

The file includes a birthweight variable.  Additonally, the one- and five-minute APGAR scores are included.  These are measures of the well being of infants just after birth.

Variable descriptions are provided as follows.

```{r}
rm(list=ls())
load("bwght_w203.RData")
desc
```


```{r}
par(mfrow=c(1,2))
hist(data$fage)
hist(data$mage)
```
```{r}
dplyr::filter(data,mage<18 | fage<18)
```
Motherage follows a fairly normal distribution. However, father age is right skewed. Might be important in analysis.
```{r}
# Simple model with only the relevant variables
df_md1 = dplyr::select(data,bwght,npvis,monpre,npvissq)
df_md1 <- na.omit(df_md1)
md1 = lm(bwght ~., data = df_md1)
summary(md1)
par(mfrow=c(2,2))
plot(md1)
par(mfrow=c(1,1))
plot(predict(md1),rstudent(md1))
```


Adjusted R_squared is 0.014011. There are polynomial relationship between
birthweight and number of visit. More visits doesn't necessarily incidate good outcome.
There are Evidence for outliers in the data in the rstudent plot.


```{r}
#best model?
df_md2 = dplyr::select(data,bwght,npvis,foth,cigs,male,fage,monpre,npvissq)
df_md2 <- na.omit(df_md2)
md2 = lm(bwght ~., data = df_md2)
summary(md2)
par(mfrow=c(2,2))
plot(md2)
par(mfrow=c(1,1))
plot(predict(md2),rstudent(md2))
```
```{r}
# problematic variable, father education
df_md3 = dplyr::select(data,bwght,npvis,foth,cigs,male,fage,monpre,npvissq,feduc)
df_md3 <- na.omit(df_md3)
md3 = lm(bwght ~., data = df_md3)
summary(md3)
par(mfrow=c(2,2))
plot(md3)
par(mfrow=c(1,1))
plot(predict(md3),rstudent(md3))
```


Adding father education reduces the adjusted R-squared. 











