---
title: "HW week 11"
subtitle: "w203: Statistics for Data Science"
author: "Ted Pham"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Get familiar with the data
You receive a data set from World Bank Development Indicators. 

  - Load the data using `load` and see what is loaded by using `ls()`. You should see `Data` which is the data frame including data, and `Descriptions` which is a data frame that includes variable names. 
  - Look at the variables, read their descriptions, and take a look at their histograms. Think about the transformations that you may need to use for these variables in the section below. 
  - Run: `apply(!is.na(Data[,-(1:2)] ) , MARGIN= 2, mean )` and explain what it is showing.
  - Can you include both `NE.IMP.GNFS.CD` and `NE.EXP.GNFS.CD` in the same OLS model? Why?
  - Rename the variable named `AG.LND.FRST.ZS` to `forest.` This is going to be our dependent variable.

```{r}
load("Week11.Rdata")
ls()
names(Data)
```

```{r}
Definitions
```

```{r}
summary(Data)
```
### Decribe a model for that predicts `forest`


      
Include all variables except country name and code in the model.
Find the 2 with the lowest p-value are and the fewest # of NA's

Agricultural raw materials exports and GDP (NY.GDP.MKTP.CD)
```{r}
lm.test = lm(AG.LND.FRST.ZS~ . -Country.Name - Country.Code,data = Data)
summary(lm.test)
```


  - Write a model with two explanatory variables. 
    - Create a residuals versus fitted values plot and assess whether your coefficients are unbiased.
```{r}
lm.fit2var = lm(AG.LND.FRST.ZS ~ TX.VAL.AGRI.ZS.UN + NY.GDP.MKTP.CD, data = Data)
par(mfrow=c(2,2))
plot(lm.fit2var)

```


```{r}
 plot(predict(lm.fit2var),rstudent(lm.fit2var))
 
```
```{r}
summary(lm.fit2var)
```

```{r}
#Check for unbiasness
#Get the rows used in the fitting 
data_use = model.frame(lm.fit2var)
print(sum(data_use$TX.VAL.AGRI.ZS.UN*residuals(lm.fit2var)))
print(sum(data_use$NY.GDP.MKTP.CD*residuals(lm.fit2var)))
```

The sum(X*u)'s are roughly equal to zero so we can assume that the estimators are unbiased.
Also by looking at the standardized residuals vs fitted, we can also see that the residuals are randomly
distributed around zero, indicative of unbiasness.


- How many observations are being used in your analysis? 
- Are the countries that are dropping out dropping out by random chance? If not, what would this do to our inference?

The number of observations used in the regression is 206.

```{r}
length(residuals(lm.fit2var))
```
```{r}
#get the list of country ommitted 
k = data_use$AG.LND.FRST.ZS
m = data_use$TX.VAL.AGRI.ZS.UN
l = subset(Data,!Data$AG.LND.FRST.ZS %in% k)
data_used = subset(Data, Data$AG.LND.FRST.ZS %in% k & Data$TX.VAL.AGRI.ZS.UN %in% m)
l$Country.Name
```
The list of countries ommitted resulted from the NA's from the three columns extracted from Data for the analysis.
These appears to be the territories of large countries or small countries without accurate reported GDP.
The number of countries omitted was minimized so it would affect minimally on our model.

```{r}
  summary(Data$MS.MIL.XPND.ZS)
```
  - Now add a third variable.
  - Show how you would use the regression anatomy formula to compute the coefficient on your third variable.  First, regress the third 
```{r}
model3var1= lm(data_used$MS.MIL.XPND.GD.ZS~lm.fit2var$residuals)
summary(model3var1)
```


variable on your first two variables and extract the residuals.  Next, regress forest on the residuals from the first stage.


```{r}
lm.fit3var = lm(AG.LND.FRST.ZS ~ TX.VAL.AGRI.ZS.UN + NY.GDP.MKTP.CD + MS.MIL.XPND.GD.ZS, data = Data)
summary(lm.fit3var)
```
  - Compare your two models.
```{r}
AIC(lm.fit3var) > AIC(model3var1)
```

- Do you see an improvement? Explain how you can tell.

Did not see an improvement because the AIC of the 2nd model is larger than the one obtained from regress the residuals.
### Make up a country

  - Make up a country named `Mediland` which has every indicator set at the median value observed in the data. 
  - How much forest would this country have?
  
  Mediland would have 30.95% forest area
  
```{r}
mediland = data.frame("TX.VAL.AGRI.ZS.UN"=1.60804,"NY.GDP.MKTP.CD" = 5.262e+10,"MS.MIL.XPND.GD.ZS" =1.535 )
predict(lm.fit3var,newdata = mediland, interval = "confidence")
```
    
### Take away

  - Agricultural raw materials exports (% of merchandise exports) is representative of % forest area.
  - Many NA's values affect the strength of multiregression and not all variables are relevant to the models in common sense.
  - Must be wary about the fitness of the data for linear regression



